---
title: "Introduction to linear regression with heterogeneity in panel data"
author: "Xingyan Xu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to linear regression with heterogeneity in panel data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

_SA24204168_ is an R package for performing grouped panel data regression (implemented through the R package _Rcpp_). The _SA24204R.R_ file contains two user-accessible functions: _dif_ and _gpd_. The purpose of _dif_ is to difference both the independent and dependent variables of the panel data along the time axis, resulting in differenced variables. This helps to effectively eliminate the impact of individual effects on parameter estimation. The purpose of _gpd_ is to fit a grouped panel data regression model, where it determines the group assignments for each panel data unit based on the specified number of groups, estimates the corresponding group values, and calculates the estimated IC value to evaluate the quality of the model’s estimates.The _SA24204C.cpp_ file contains a function count, which calculates the loss function. By inputting the panel data x, its corresponding response variable y, the estimated intercept value beta, and the specific panel unit index, the function returns the loss function value for the i-th panel unit.

## model
In practical regression tasks, we often encounter situations where two groups of regression data with different slopes are mixed together, such as:
```{r, warning = FALSE}
# 加载必要的包
library(ggplot2)

# 设置随机数种子，确保可重现性
set.seed(42)

# 生成数据
n <- 20  # 每组20个数据

# 第一组数据: y = x + epsilon
x1 <- runif(n, 0, 1)  # 自变量 x 服从 U(0, 1)
epsilon1 <- rnorm(n, mean = 0, sd = 0.4)  # 噪声 epsilon 服从 N(0, 0.4)
y1 <- x1 + epsilon1  # 计算响应变量 y

# 第二组数据: y = 3x + epsilon
x2 <- runif(n, 0, 1)  # 自变量 x 服从 U(0, 1)
epsilon2 <- rnorm(n, mean = 0, sd = 0.4)  # 噪声 epsilon 服从 N(0, 0.4)
y2 <- 5 * x2 + epsilon2  # 计算响应变量 y

# 合并数据
data <- data.frame(
  x = c(x1, x2),
  y = c(y1, y2),
  group = factor(rep(1:2, each = n))  # 分组标签
)

# 做回归分析
model1 <- lm(y ~ x, data = data[data$group == 1, ])
model2 <- lm(y ~ x, data = data[data$group == 2, ])

# 整体回归模型 (使用所有数据)
model_all <- lm(y ~ x, data = data)

# 画回归图
ggplot(data, aes(x = x, y = y, color = group)) +
  geom_point() +  # 绘制散点
  geom_smooth(method = "lm", se = FALSE) +  # 绘制每组的回归曲线
  geom_smooth(data = data, aes(x = x, y = y), method = "lm", color = "black", linetype = "dashed", se = FALSE) +  # 整体回归线，黑色虚线
  labs(title = "Regression Lines for Two Groups with Overall Fit",
       x = "x",
       y = "y") +
  scale_color_manual(values = c("blue", "red")) +  # 设置不同组别的颜色
  theme_minimal()

```

In the plot, the **blue line** represents the linear regression curve fitted to the data generated from the model \( y = x + \epsilon \), 
where \( \epsilon \sim N(0, 0.4) \). The **red line** represents the linear regression curve fitted to the data generated from the model 
\( y = 5x + \epsilon \). The **black dashed line** represents the linear regression curve fitted to all the data points combined.

As we can observe, if we fit the data without distinguishing the two groups, the slope of the fitted line (black dashed line) differs significantly from the true value of 1 (from the first group) and 5 (from the second group). This demonstrates that the overall regression model fails to capture the different behaviors in the two groups, leading to poor model fit.

We primarily consider the following model:

$$
y_{it} = \alpha_{i} + x_{it}\beta_{g_i} + \epsilon_{it}, \quad \epsilon_{it} \overset{\text{iid}}{\sim} \, N(0, \sigma^2)
$$

where \( i = 1, \dots, n \) represents the number of panel data units,  \( t = 1, \dots, T \) represents the time periods， and $\beta_{g_i}$ represents the i-th unit belongs to group $g_i$, where $g_i=1,2,..,G$, G is the number of groups.

Therefore, we need to identify the group each sample unit belongs to and determine the slope for each group.

## R function _dif_ and _gpd_

Input and output of _dif_ are as follows:
```{r}
function(x, y) {
  # Input: Panel scalar data x and its corresponding response variable y.
  # Output: The differenced data x_diff and y_diff along the time axis.

  n_y <- nrow(y)  
  t_y <- ncol(y)  
  
  n_x <- nrow(x)  
  t_x <- ncol(x)  
  
  if (n_y != n_x) {
    stop("The number of rows (n) in x and y must be the same.")
  }
  
  if (t_y != t_x) {
    stop("The number of columns (t) in x and y must be the same.")
  }
  
  if (t_y <= 1) {
    stop("The number of columns (t) in both x and y must be greater than 1.")
  }
  
  y_diff <- data.frame(matrix(NA, nrow = n_y, ncol = t_y - 1)) 
  x_diff <- data.frame(matrix(NA, nrow = n_x, ncol = t_x - 1))  

  for (i in 1:n_y) {
    for (j in 2:t_y) { 
      y_diff[i, j - 1] <- y[i, j] - y[i, j - 1] 
      x_diff[i, j - 1] <- x[i, j] - x[i, j - 1] 
    }
  }
  
  colnames(y_diff) <- paste("Y", 1:(t_y - 1), sep = "")  
  colnames(x_diff) <- paste("X", 1:(t_x - 1), sep = "")  
  
  return(list(y_diff = y_diff, x_diff = x_diff))
}
```

**Note**: The input matrices **x** and **y** are both of size **n × t**, where **x[i, ]** and **y[i, ]** represent the independent and dependent variables of the **i-th** unit in the panel data over time. On the other hand, **x[, j]** and **y[, j]** represent the cross-sectional data of the **j-th** time point. Thus, **$x[i, t]** and **y[i, t]** are the independent and dependent variables of the **i-th** unit at the **j-th** time point, and both are scalars.

The purpose of the `_dif_` function is to eliminate the influence of the fixed effect $\alpha_i$ in the model.

Then we focus on the function `_gpd_`. Due to the length of the function, we will not display the full code. The purpose of the `_gpd_` function is to take panel data `x` and `y`, along with the number of groups `G`, and output the group assignment for each panel data unit as well as the estimated slope for each group. It also computes the IC value to assess the goodness of fit of the estimated model. Specifically, the smaller the IC value, the better the model fit. The value of `G` that minimizes the IC value is considered the best estimate for the number of groups.

## C++ function _count_
```{r,eval=FALSE}
double count(NumericMatrix x, NumericMatrix y, double beta, int i) {
  // 初始化平方误差的和
  double sum = 0.0;
  
  // 获取 y 和 x 对应行的差异 (y[i, ] - x[i, ] * beta)
  for (int j = 0; j < x.ncol(); j++) {
    sum += pow(y(i, j) - x(i, j) * beta, 2);  // 计算每个元素的差的平方并累加
  }
  
  return sum;
}
```

The function `_count_` is responsible for calculating the loss function for the \(i\)-th unit. The loss function is in the form of the least squares loss.
